---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(magrittr)
get_read <- function(link){
    if(!file.exists('activity.zip')){
        download.file(link, 'activity.zip', method = 'curl')
    }
    
    if(!file.exists('activity.csv')){
        unzip('activity.zip')
    }
    
    read.csv('activity.csv')
}    

process <- function(data){
    data$date <- as.Date(data$date)
    return(data)
}

activity_data <- get_read('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip') %>%
    process()
```


## What is mean total number of steps taken per day?
```{r message=FALSE}
library (dplyr)

# Create data frame of daily total, mean, and median steps
per_day <- function(data){
    group_by(data, date) %>%
        summarise(sum(steps,na.rm = TRUE))
}

# Call funtion to build df and set column names
daily_steps <- per_day(activity_data)
names(daily_steps) <- c('date', 'total_steps')

# Create histogram of total daily steps
hist(daily_steps$total_steps, xlab= 'Date', ylab = 'Steps', main = 'Daily total steps')

```

Mean:
```{r}
mean(daily_steps$total_steps)
```
Median:
```{r}
median(daily_steps$total_steps)
```


## What is the average daily activity pattern?
```{r}
# Create data frame of activity mean by interval
avg_day <- function(data){
    group_by(data, interval) %>%
        summarise(mean(steps,na.rm = TRUE))
}

# Call function to get interval mean and name columns
all_day <- avg_day(activity_data) 
names(all_day) <- c('interval', 'mean_steps')

# Create the plot
with(all_day,
     plot(interval,mean_steps, 
          xlab="Interval", ylab= "Steps", 
          main="Average Daily Activity Pattern", 
          type="l"))

# Find the max activity and add it to the chart
max_active <- all_day[all_day$mean_steps ==
                          max(all_day$mean_steps),]$interval
abline(v = max_active,col="blue")
```

## Imputing missing values

```{r}
# Copy base data to work with
filled <- activity_data

# boolian for if steps are na (they are illmatic)
nas <- is.na(filled$steps)

# Get number of rows with missing values
sum(nas)
```

I decided that the most acurate way to fill the data would be to fill in each interval with the mean for intervals calulated earlier. This should give us a good representation of a normal day/interval when data is missing.
```{r}
# Replace NA values with the matching interval mean calculated earlier
filled[nas,]$steps <- 
    match(filled[nas,]$interval, all_day$interval) %>%
    all_day$mean_steps[.]

# Call function from above to get sum, mean, and median per day
filled_daily<-per_day(filled)
names(filled_daily) <- c('date', 'total_steps')

# Create histogram of total daily steps
hist(filled_daily$total_steps, xlab= 'Date', ylab = 'Steps', main = "Daily Steps after filling data")
```

Mean:
```{r}
mean(filled_daily$total_steps)
```
Median:
```{r}
median(filled_daily$total_steps)
```

The histogram is mostly affected by transfering a lot of days with NAs which were in the 0-5000 steps section to the most common 10,000 to 15,000 section. 

The mean and median are both increased which makes sense with the daily totals being larger from filling in the missing data.  

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(ggplot2)

filled$weekend <- 
    weekdays(filled$date) == "Sunday" | 
    weekdays(filled$date) == "Saturday"

filled$wkdy_label <- ifelse(filled$weekend, "Weekend", "Weekday")

avg_day_wkend <- function(data){
    group_by(data, interval, wkdy_label) %>%
        summarise(mean(steps,na.rm = TRUE))
}

wk_wkend <- avg_day_wkend(filled)
names(wk_wkend)[3] <- 'steps'

qplot(interval, steps, geom = "line", facets = wkdy_label ~ . , data= wk_wkend)

```

